---
title: "Tree Ring Plotting"
author: "Kelly Heilman"
date: "Monday, May 04, 2015"
output: html_document
---


I read in data from White oaks across the prairie peninsula from ITRDB
```{r}
library(dplR)
filename_list <- c("in001.rwl", "in002.rwl",
             "ia001.rwl","ia002.rwl", "ia003.rwl",
                   "ia004.rwl","ia005.rwl","ia006.rwl", 
                   "ia012.rwl", "ia015.rwl","ia016.rwl", "ia018.rwl","ia020.rwl","ia021.rwl", "ia023.rwl", "ia024.rwl","ia025.rwl","ia026.rwl","ia027.rwl", "ia032.rwl", "ia033.rwl","ia034.rwl","il001.rwl", "il002.rwl","il008.rwl","il009.rwl","il010.rwl", "il011.rwl","il012.rwl","il013.rwl", "il014.rwl","ar002.rwl","ar009.rwl","ar013.rwl", "ar014.rwl","ar015.rwl","ar018.rwl","ar059.rwl","ar062.rwl", "mo004.rwl","mo002.rwl", "mo007.rwl","mn022.rwl", "mn023.rwl","mn028.rwl","mn029.rwl","mn030.rwl","mn016.rwl","oh001.rwl","oh002.rwl", 'oh003.rwl',"oh006.rwl", "mi005.rwl", "mi013.rwl", "mi018.rwl",
"ky002.rwl", "ky004.rwl", "ks012.rwl",
"wi006.rwl","wi008.rwl", "wi009.rwl","wi010.rwl","tn005.rwl","tn008.rwl","tn017.rwl","tn020.rwl","tn021.rwl","sd003.rwl", "sd006.rwl", "sd007.rwl", "sd009.rwl", "sd010.rwl", "sd011.rwl", "sd013.rwl", "sd014.rwl", "sd015.rwl", "sd018.rwl", "sd019.rwl","sd020.rwl","sd021.rwl","sd022.rwl","sd023.rwl","sd024.rwl")
destfile=filename_list

#tried l_ply first, but this loops seems to work better
###already downloaded these files, so they are commented out:

for(i in 1:length(filename_list)){
 download.file(filename_list[i], url=paste0('ftp://ftp.ncdc.noaa.gov/pub/data/paleo/treering/measurements/northamerica/usa/', filename_list[i]), method="auto", quiet = FALSE, mode = "w",cacheOK = TRUE)
}

#filenames <- list.files("C:/Users/Kelly/Documents/TreeRings", pattern="*.rwl", full.names=FALSE)
#filenames<-filenames[2:93]
filenames<- filename_list
myDir <- "C:/Users/Kelly/Documents/TreeRings" 

data_names <- gsub("[.]rwl", "", filenames) 

for(i in 1:length(filenames)) assign(data_names[i], read.rwl(file.path(myDir, filenames[i])))

```
Detrend the rwl data
```{r}
obj<- list(data_names)
#trying to detrend all the data

#make plots of each detrending and spline
pdf("detrending_splines.pdf")
for(i in 1:length(data_names)) {
  nam <- paste0(data_names[i], ".rwi")
  assign(nam, detrend(rwl=get(data_names[i]), y.name = colnames(get(data_names[i])), make.plot = TRUE,method = "Spline" ))}
dev.off()

```

Make site level chronologies and plot these rwi over time
```{r}
#make a list of detrended objects
rwis <- objects(pattern="*.rwi")

#create chronologies for the detrended objects
for(i in 1:length(rwis)) {
  nam <- paste0(rwis[i], ".chron")
  assign(nam, chron(x=get(rwis[i]), prefix = substr(colnames(get(rwis[i]))[1],1,3)))}

#plot each chronology
chrons<- objects(pattern="*.rwi.chron")
pdf("individual.chron.plots.pdf")
for(i in 1:length(chrons)){
plot(get(chrons[i]),xlab="Year",ylab="RWI", main=chrons[i])
}
dev.off()
```
########
left of here on round 2 with a larger number of itrdb sites

Are these RWI trends correlated to zonal index?
```{r}

library(plyr)
library(reshape)
library(ggplot2)

sites<- read.csv("itrdb_locs.csv", header=TRUE, stringsAsFactors=FALSE)

#create a column in sites that matches the names of our chronologies
sites$chron.code <- paste0(sites$ITRDB.code, ".rwi.chron") 

zonal.index <- read.csv("C:/Users/Kelly/Documents/Undergrad_projects/clim-analysis/geopthgt1980_2013.csv")
zonal.index$years<- 1977:2013

zonal.index <- zonal.index[3:nrow(zonal.index),]

for(i in 1:length(chrons)){
  nam <- paste0(chrons[i], ".mod")
  assign(nam, merge(get(chrons[i]),zonal.index, by.x="row.names", by.y="years"))
}

files<- rbind.fill(ia001.rwi.chron.mod,ia002.rwi.chron.mod,ia003.rwi.chron.mod,ia004.rwi.chron.mod,ia005.rwi.chron.mod ,ia012.rwi.chron.mod, ia015.rwi.chron.mod,ia021.rwi.chron.mod, ia023.rwi.chron.mod,ia024.rwi.chron.mod , ia026.rwi.chron.mod,ia030.rwi.chron.mod,il001.rwi.chron.mod, il008.rwi.chron.mod, il009.rwi.chron.mod ,il010.rwi.chron.mod ,il011.rwi.chron.mod,il013.rwi.chron.mod,il014.rwi.chron.mod,in001.rwi.chron.mod,mn022.rwi.chron.mod , mo002.rwi.chron.mod ,mo004.rwi.chron.mod, mo007.rwi.chron.mod,mo015.rwi.chron.mod,mo018.rwi.chron.mod, mo050.rwi.chron.mod, wi006.rwi.chron.mod )    

IA <- c("BKBstd", "NEGstd", "LKSstd", "GEOstd", "PKPstd", "783std", "782std", "788std", "648std", "464std", "649std", "WPHstd")
IL <- c("SAWstd", "STRstd","KANstd", "LNSstd", "FXRstd", "FCLstd", "GCYstd")
IN <-c("PW1std")
MN <-c("SBOstd")
MO <-c("K1Astd", "K3Astd", "K4Astd", "WEGstd", "RORstd", "CRNstd")
WI <-c("PlPstd")

df.new<-melt(files, id=c("Row.names", "samp.depth", "X", "east.zonal.index", "hemispheric.zonal.index", "west.zonal.index"))
data.frame(df.new)
ia.trues <- df.new$variable %in% IA
il.trues <- df.new$variable %in% IL
in.trues <- df.new$variable %in% IN
mn.trues <- df.new$variable %in% MN
mo.trues <- df.new$variable %in% MO
wi.trues <- df.new$variable %in% WI

#create new column by group
df.new$group<-"NA"

df.new[ia.trues,]$group <- "IA"
df.new[il.trues,]$group <- "IL"
df.new[in.trues,]$group <- "IN"
df.new[mn.trues,]$group <- "MN"
df.new[mo.trues,]$group <- "MO"
df.new[wi.trues,]$group <- "WI"

lats.data <- read.csv("itrdb_locs.csv")

data.locs<-merge(df.new, lats.data, by.x ="variable", by.y="chroncd")

plot.E.zonal <- ggplot(data.locs, aes( east.zonal.index, value))+
  geom_point(aes(color=group, size=3))

plot.lat <- ggplot(data.locs, aes(lat, value))+
  geom_point(aes(color=group, size=3))

plot.long <- ggplot(data.locs, aes(lon, value))+
  geom_point(aes(color=lon, size=3))
      
```
There does not appear to be a strong trend in correlation of ring width and zonal index data

Are there latitudinal/longitudinal gradients in the data?
Extract precipitation data and sum for each year to get annual precipitation
```{r}
#download NOAA data, save in directory
library(raster)

locs <- read.csv("itrdb_locs.csv")
coordinates(locs) <-~lon+lat
proj4string(locs)<- "+proj=longlat +ellps=WGS84 +datum=WGS84 +towgs84=0,0,0"

library(ncdf4)
library(raster)
inputfile="C:/Users/Kelly/Documents/TreeRings/precip.mon.total.v6.nc"
p<-nc_open(inputfile)
h<-inputfile
pr<-brick(h,lvar=3, varname="precip")

inputfile="C:/Users/Kelly/Documents/TreeRings/air.mon.mean.v301.nc"
p<-nc_open(inputfile)
h<-inputfile
tmean<-brick(h,lvar=3, varname="air")



locs <- read.csv("itrdb_locs.csv")
locs$long<-locs$lon+360
locs.soil<-locs
coordinates(locs) <-~long+lat
coordinates(locs.soil)<- ~lon+lat
proj4string(locs)<- "+proj=longlat +ellps=WGS84 +datum=WGS84 +towgs84=0,0,0"
proj4string(locs.soil)<- "+proj=longlat +ellps=WGS84 +datum=WGS84 +towgs84=0,0,0"

#these soil layers will be used much later
#clay raster
clay <- raster("soils_data/clay_rast1.tif")

#sand raster
sand <- raster("soils_data/sand_rast1.tif")

#silt raster
silt <- raster("soils_data/silt_rast1.tif")
#awc raster
awc <- raster("soils_data/awc_rast1.tif")
#elev raster
elev <- raster("soils_data/elev_rast1.tif")
#ksat
ksat <- raster("soils_data/ksat_raster1.tif")
#slope raster
drainage <- raster("soils_data/drain_rast1.tif")

#now extract point level data at each location
pr.data <- extract(pr, locs)
temp.data <- extract(tmean, locs)

#extract soil data from locs.soil and add to locs
clay.data <- extract(clay, locs.soil)
locs$clay <- clay.data

sand.data <- extract(sand, locs.soil)
locs$sand <- sand.data

silt.data <- extract(silt, locs.soil)
locs$silt <- silt.data

awc.data <- extract(awc, locs.soil)
locs$awc <- awc.data

ksat.data <- extract(ksat, locs.soil)
locs$ksat <- ksat.data

drainage.data <- extract(drainage, locs.soil)
locs$drainage <- drainage.data

elev.data <- extract(elev, locs.soil)
locs$elev <- elev.data
#pr.data now has all the months from 1901-2011

#read in the itrdb_locs again so it does not have the soil factors
locs.1 <- read.csv("itrdb_locs.csv")
#now need to sum every 12 values to get the total preciptiation each year
pr.data.1<-as.data.frame(pr.data)
pr.data.1[29, 1:1320]<- group #why do i need group?
pr.df<-t(pr.data.1)
pr.df<-data.frame(pr.df)
pr.df$year <- rep(1901:2010, each=12) #create year index
codes<- locs.1$ITRDB.code
#but using data.table to sum by year works
dt <- data.table(pr.df)
yearly.precip<-dt[,lapply(.SD, sum) ,by = year]
#convert back to data.frame
yearly.precip<-data.frame(yearly.precip)
colnames(yearly.precip)<-c("year", as.character(codes))


#averaging the temperature data
temp.data.1<-as.data.frame(temp.data)
#temp.data.1[29, 1:1320]<- group #why do i need group?
temp.df<-t(temp.data.1)
temp.df<-data.frame(temp.df)
#temperature data goes from 1900 to 2010
temp.df$year <- rep(1900:2010, each=12) #create year index
codes<- locs.1$ITRDB.code

#but using data.table to sum by year works
dt <- data.table(temp.df)
yearly.temp<-dt[,lapply(.SD, mean) ,by = "year"]
#convert back to data.frame
yearly.temp<-data.frame(yearly.temp)
codes<- locs.1$ITRDB.code
colnames(yearly.temp)<-c("year", as.character(codes))

yr.data<-t(yearly.precip)
yr.data.no.yr<- data.frame(yr.data[2:29,])
yr.data.no.yr$chroncd<- locs.1$chroncd
yr.data.no.yr$site<-locs.1$ITRDB.code

yr.data.t<-t(yearly.temp)
yr.data.no.yr.t<- data.frame(yr.data.t[2:29,])
#add chron code column
yr.data.no.yr.t$chroncd<- locs$chroncd
#add cite code column
yr.data.no.yr.t$site<-locs.1$ITRDB.code

#merge temperature data with locs
locs.temp <- merge(data.frame(locs.1), data.frame(yr.data.no.yr.t), by='chroncd')
#merge the product of above with locs
locs.pr <- merge(data.frame(locs.1), data.frame(yr.data.no.yr), by='chroncd')
## it is probably best to create a function to merge the chron to each site and plot
locs.pr.t<-t(locs.pr)
names<-as.character(locs.pr.t[1,])
locs.pr.t<-locs.pr.t[6:115,]
locs.pr.t<-data.frame(locs.pr.t)
colnames(locs.pr.t)<-names
locs.pr.t$year<- as.character(1901:2010)

locs.temp.t <-t(locs.temp)
names <- locs.temp.t[1,]
locs.temp.t<-locs.temp.t[7:117,]
locs.temp.t<-data.frame(locs.temp.t)
colnames(locs.temp.t) <- names
locs.temp.t$year<- as.character(1900:2010)

#objects that have rwi.chron
rwi.chrons <- ls(pattern = '\\.rwi.chron$')

#this gross for loop merges the location temperature data with the rwi.chronology for each location, plots temp vs. rwi for all locations in separate plots, and saves the correlation coeffiicient for each site in a new object
newobj = list()
corobj = list()

pdf("temperature_rwi_plots.pdf")
for(i in 1:length(rwi.chrons)){
newobj[[i]] <- merge(data.frame(locs.temp.t[,c(colnames(get(rwi.chrons[i])[1]), 'year')]), data.frame(get(rwi.chrons[i])), by.x='year', by.y=0)

#plot each chronology against its site temperature
plot(t(newobj[[i]][2]), t(newobj[[i]][3]), xlab="Tmean (C)", ylab="RWI", main=paste(colnames(get(rwi.chrons[i])[1]),"Chronology", sep="" ))

#abline(lm(data.frame(unlist(newobj[[i]][3])) ~ data.frame(unlist(newobj[[i]][2])))

assign(paste0(colnames(get(rwi.chrons[i])[1]),".temp", sep = ""), newobj[[i]])

corobj[[i]] <- cor(data.frame(as.numeric(t(newobj[[i]][2]))), data.frame(as.numeric(t(newobj[[i]][3]))))

assign(paste0(colnames(get(rwi.chrons[i])[1]),".corr.temp", sep = ""), corobj[[i]])
}
dev.off()


#do the same for precipitation plots
newobject<- list()
corobject<- list()

pdf("precipitation_rwi_plots.pdf")
for(i in 1:length(rwi.chrons)){
  
newobject[[i]] <- merge(data.frame(locs.pr.t[,c(colnames(get(rwi.chrons[[i]])[1]),"year")]), data.frame(get(rwi.chrons[i])), by.x="year", by.y=0)

#plot each chronology against its site temperature
plot(t(newobject[[i]][2]), t(newobject[[i]][3]), xlab="Annual Precip. (mm)", ylab="RWI", main=paste(colnames(get(rwi.chrons[i])[1]),"Chronology", sep="" ))

#abline(lm(as.vector(get(newobj[[i]][3])) ~ as.vector(get(newobj[[i]][2])))

assign(paste0(colnames(get(rwi.chrons[i])[1]),".pr", sep = ""), newobject[[i]])

corobject[[i]] <- cor(data.frame(as.numeric(t(newobject[[i]][2]))), data.frame(as.numeric(t(newobject[[i]][3]))))

assign(paste0(colnames(get(rwi.chrons[i])[1]),".corr.precip", sep = ""), corobject[[i]])
}
dev.off()



###may not need these functions, now that we have above

#this is the same function above, but with temperature dat
merge.chrons.t<- function(location.rwi.chron){
newobj <- merge(data.frame(locs.temp.t[,c(colnames(location.rwi.chron)[1] , 'year')]), data.frame(location.rwi.chron), by.x='year', by.y=0)
colnames(newobj) <- c("year", "temp", "rwi", "samp.depth")
return(newobj)
}

#lapply over list of rwi.chrons
#doesnt work here
lapply(get(rwi.chrons), FUN= merge.chrons.t)

plot.m <- function(object.column, object.column2){
  plot(as.numeric(as.vector(object.column), object.column2))
  lm(object.column2 ~ as.numeric)
}

#melt function for .rwi files and locs precip
melt.rwi<- function(location.rwi) {data.frame(melt(merge(data.frame(locs.pr.t), data.frame(location.rwi), by.x='year', by.y=0), value.name = "year"))
}

#melt function for .rwi files and locs temp
melt.rwi.temp<- function(location.rwi) {data.frame(melt(merge(data.frame(locs.temp.t), data.frame(location.rwi), by.x='year', by.y=0), value.name = "year"))
}

```
This section plots correlation coefficients with latitude & longitude
```{r}
#get all the corr.precip objects and put in a dataframe
corr.precips <- ls(pattern = '\\corr.precip$')
names<- corr.precips
for(i in 1:length(corr.precips)){
  dat[i] <- get(corr.precips[[i]])
  chroncd[i] <- strsplit(names[i], ".corr.precip", 6)
  }

pr.corr<- data.frame(corr=dat,
          chroncd=unlist(chroncd), 
          stringsAsFactors=FALSE)

#do the same for temperature data
corr.temps <- ls(pattern = '\\corr.temp$')
names<- corr.temps
dat.t<- dat
chroncd.t<- chroncd
for(i in 1:length(corr.temps)){
  dat.t[i] <- get(corr.temps[[i]])
  chroncd.t[i] <- strsplit(names[i], ".corr.temp", 6)
  }

temp.corr<- data.frame(corr=dat.t,
          chroncd=unlist(chroncd.t), 
          stringsAsFactors=FALSE)


#merge temp.corr and pr.corr dfs with location df to get lat long values
prcorr.df <- merge(pr.corr, locs, by="chroncd")
tempcorr.df <- merge(temp.corr, locs, by="chroncd")

pdf("outputs/correlation_lat_lon.pdf")
#plot the relationship with latitude
plot(prcorr.df$lat, prcorr.df$corr, pch=15, xlab="latitude", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across Latitudes")
lm.copr <- lm(prcorr.df$corr ~ prcorr.df$lat)
abline(lm.copr)

plot(tempcorr.df$lat, tempcorr.df$corr, pch=15, xlab="latitude", ylab="Correlation coefficient (RWI vs. Mean Temperature)", main="Temperature Correlation coefficient of tree ring width index across Latitudes")
lm.cotemp <- lm(tempcorr.df$corr ~ tempcorr.df$lat)
abline(lm.cotemp)

library(rgdal)
#make the precipitation data spatial and plot on the map
coordinates(prcorr.df)<-~long+lat
proj4string(prcorr.df) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +towgs84=0,0,0"
prcorr.NAD83 <- spTransform(prcorr.df, CRS("+proj=longlat +datum=NAD83"))
prcorr.df <- data.frame(prcorr.NAD83)

#make the temperature data spatial and plot on map
coordinates(tempcorr.df)<-~long+lat
proj4string(tempcorr.df) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +towgs84=0,0,0"
tempcorr.NAD83 <- spTransform(tempcorr.df, CRS("+proj=longlat +datum=NAD83"))
tempcorr.df <- data.frame(tempcorr.NAD83)

#load in map outline of states 
library(maps)
#load us map data
all_states <- map_data("state")
states <- subset(all_states, region %in% c( "indiana", "illinois", "missouri", "iowa", "wisconsin", "minnesota"  ) )
coordinates(states)<-~long+lat
class(states)
proj4string(states) <-CRS("+proj=longlat +datum=NAD83")
mapdata<-states
#mapdata<-spTransform(states, CRS("+proj=aea +lat_1=42.122774 +lat_2=49.01518 +lat_0=45.568977 +lon_0=-83.248627 +x_0=1000000 +y_0=1000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"))
mapdata<-data.frame(mapdata) #ggplot needs it in a data.frame


#plot the precipitation points on the map
p <-ggplot() +geom_polygon(data=mapdata, aes(x=long, y=lat, group = group))
p <- p + geom_point( data=prcorr.df, aes(x=lon, y=lat, color = corr, size = corr)) +
  scale_colour_gradient(low = "red") + 
  labs(x="easting", y="northing", title="Correlation coefficient of tree ring width index to Annual precipitaiton ") 
p

#plot the temp points on map
t <-ggplot() +geom_polygon(data=mapdata, aes(x=long, y=lat, group = group))
t <- t + geom_point( data=tempcorr.df, aes(x=lon, y=lat, color = corr, size = abs(corr))) +
  scale_colour_gradient(low = "red") + 
  labs(x="easting", y="northing", title="Correlation coefficient of tree ring width index to Mean annual temperature ") 
t
dev.off()

```

How does soil type play into these trends?
```{r}

#lat.plot<-ggplot(prcorr.df,aes(lat, prcorrs))+geom_point(col="blue") + stat_smooth(method="lm", se=FALSE)+ labs(x="latitude", y="Correlation coefficient (RWI vs. Annual Precip)", title="Precip Correlation coefficient of tree ring width index across Latitudes") 

#long.plot<-ggplot(prcorr.df,aes(lon, prcorrs))+geom_point(col="blue") + stat_smooth(method="lm", se=FALSE)+ scale_y_continuous(limits = c(0, 0.5))+labs(x="longitude", y="Correlation coefficient (RWI vs. Annual Precip)", title="Precip Correlation coefficient of tree ring width index across Longitudes") 
pdf("outputs/soil_correlation.pdf")
#plot the relationship with longitude
plot(prcorr.df$lon, prcorr.df$corr, pch= 15,xlab="longitude", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across Longitudes")

lm.coprlon <- lm(prcorr.df$corr ~ prcorr.df$lon)
abline(lm.coprlon)

#sand is not significant
plot(prcorr.df$sand, prcorr.df$corr, pch=15, xlab="% sand", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across Sand gradients")
lm.sand <- lm(prcorr.df$corr ~ prcorr.df$sand)
abline(lm.sand)

#silt is not significant
plot(prcorr.df$silt, prcorr.df$corr, pch=15, xlab="% silt", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across Silt gradients")
lm.silt <- lm(prcorr.df$corr ~ prcorr.df$silt)
abline(lm.silt)

#clay is not significant
plot(prcorr.df$clay, prcorr.df$corr, pch=15, xlab="% clay", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across clay gradients")
lm.clay <- lm(prcorr.df$corr ~ prcorr.df$clay)
abline(lm.clay)

#awc is not significant
plot(prcorr.df$awc, prcorr.df$corr, pch=15, xlab="AWC", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across awc gradients")
lm.awc <- lm(prcorr.df$corr ~ prcorr.df$awc)
abline(lm.awc)

#ksat not significant (p = 0.1911)
plot(prcorr.df$ksat, prcorr.df$corr, pch=15, xlab="ksat", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across ksat gradients")
lm.ksat<- lm(prcorr.df$corr ~ prcorr.df$ksat)
abline(lm.ksat)

#elev not significant (p = 0.614)
plot(prcorr.df$elev, prcorr.df$corr, pch=15, xlab="elev", ylab="Correlation coefficient (RWI vs. Annual Precip)", main="Precip Correlation coefficient of tree ring width index across elev gradients")
lm.elev<- lm(prcorr.df$corr ~ prcorr.df$elev)
abline(lm.elev)


coordinates(prcorr.df)<-~long+lat
proj4string(prcorr.df) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +towgs84=0,0,0"
prcorr.NAD83 <- spTransform(prcorr.df, CRS("+proj=longlat +datum=NAD83"))
prcorr.df <- data.frame(prcorr.NAD83)

#load in map 
library(maps)
#load us map data
all_states <- map_data("state")
states <- subset(all_states, region %in% c( "indiana", "illinois", "missouri", "iowa", "wisconsin", "minnesota"  ) )
coordinates(states)<-~long+lat
proj4string(states) <-CRS("+proj=longlat +datum=NAD83")
mapdata<-states
mapdata<-data.frame(mapdata)

#plot the points on the map with silt as color
p <-ggplot() +geom_polygon(data=mapdata, aes(x=long, y=lat, group = group))
p <- p + geom_point( data=prcorr.df, aes(x=lon, y=lat, color = silt, size = corr)) +
  scale_colour_gradient(low = "red") + 
  labs(x="easting", y="northing", title="Correlation coefficient of tree ring width index to Annual precipitaiton ") 
p

dev.off()

```
