---
title: "BAI_models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)

# some initial data checking:
full.ghcn <- read.csv("outputs/full.ghcn.csv")
BAI <- full.ghcn$RWI
hist(BAI)
logBAI<- hist(log(BAI)) # approximates normal when log transformed

```

## Initial bayesian linear regression of climate on Basal Area Index
# this model assumes all trees across all species + sites + forest types have the same relationship with PDSI

```{r}



summary(full.ghcn)
# get all records that have all RWI and don't have negative diams or NA diams
full.ghcn <- full.ghcn[!is.na(full.ghcn$RWI) & full.ghcn$DBH.x > 0 & !is.na(full.ghcn$DBH.x),]
Y <- as.vector(log(full.ghcn$RWI)) 
#Y     <- 100*dat[,2]
#Y     <- (Y-mean(Y))/sd(Y)

# standardise predictor variables to have mean 0 and sd = 1
DI.scaled = scale(full.ghcn$JJA.pdsi, center= TRUE, scale=TRUE)
T.scaled = scale(full.ghcn$JUNTmax, center= TRUE, scale=TRUE)
full.ghcn$DI.scaled = as.vector(scale(full.ghcn$JJA.pdsi, center = TRUE, scale = TRUE))


#DI <- as.vector( full.ghcn$JJA.pdsi )
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x
full.ghcn$DBH.scaled = as.vector(scale(full.ghcn$DBH.x, center = TRUE, scale = TRUE))
site <- full.ghcn$site
SpecCode <- full.ghcn$SpecCode

summary(lm(Y ~ log(DBH)))

summary(lm(Y ~ log(DBH) + log((pdsi + 10)/20) + log(JunTmax)  ))

# population model for the response of each BAI to each year of climate:
population_model <- "model{

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[i], inv.var) # where Yi is already log transformed

# function g()
gfunc[i] <- beta1 +  beta2*DI.scaled[i] + beta3*DBH.scaled[i] # use Drought index

res[i] <- Y[i] - gfunc[i]   
emp.new[i] ~ dnorm(gfunc[i], inv.var)
res.new[i] <- emp.new[i] - gfunc[i]

}

# Assume normal priors for betas (because they could be negative or positive)

beta1 ~ dnorm(0, 0.01)
beta2 ~ dnorm(0, 0.01)
beta3 ~ dnorm(0,0.01)



# Non-informative Prior for the inverse variances
inv.var   ~ dgamma(0.01, 0.01)
sigma     <- 1/sqrt(inv.var)

#Derived parameters
  fit <- sum(res[])
  fit.new <- sum(res.new[])



}"


sum.lin.reg <- lm(Y ~ full.ghcn$DI.scaled + full.ghcn$DBH.scaled)
lin.rgr.ests<- summary(sum.lin.reg)$coefficients
# specify initial conditions for the three chains
initsList <- list(
  list(
    # chain 1 has low estimates
    beta1 = lin.rgr.ests[1,1] - 5*lin.rgr.ests[1,2],
    beta2 = lin.rgr.ests[2,1] - 5*lin.rgr.ests[2,2], 
    
    inv.var = 0.0001
  ), 
  # chain 2 has high values
  list(
    beta1 = lin.rgr.ests[1,1] + 5*lin.rgr.ests[1,2],
    beta2 = lin.rgr.ests[2,1] + 5*lin.rgr.ests[2,2], 
    #beta3 = lin.rgr.ests[3,1] + 5*lin.rgr.ests[3,2], 
    inv.var = 1
  ), 
  list(
    # chain 3 has a mix
    beta1 = lin.rgr.ests[1,1] + 5*lin.rgr.ests[1,2],
    beta2 = lin.rgr.ests[2,1] - 5*lin.rgr.ests[2,2], 
    #beta3 = lin.rgr.ests[3,1] + 5*lin.rgr.ests[3,2], 
   inv.var = 100
  )
)

# now 
reg.model.age <- jags.model(textConnection(population_model), 
                    data = list(Y=Y, n=n, DI.scaled = full.ghcn$DI.scaled, DBH.scaled = full.ghcn$DBH.scaled), inits = initsList, n.chains = 3)

update(reg.model.age, 1000); # Burnin for 1000 samples to start, then go higher later

samp <- coda.samples(reg.model.age, 
                     variable.names=c("beta1", "beta2","sigma", "fit", "fit.new"), 
                    n.chains = 3, n.iter=2000)

summary(samp)
gelman.diag(samp)
gelman.plot(samp)


data = list(Y=Y, n=n, DI.scaled = full.ghcn$DI.scaled, DBH.scaled = full.ghcn$DBH.scaled)
inits = initsList
n.chains = 3
out.jags <- jags(data = data,
            inits = NULL,
            parameters.to.save = c("beta1", "beta2","sigma", "fit", "fit.new"),
            model.file = population_model,
            n.chains = 3,
            n.adapt = 100,
            n.iter = 2000,
            n.burnin = 1000,
            n.thin = 1)

#Examine output summary

out.jags
# posterior predictive checks:
# Sample m draws of theta from estimated posterior p(theta|y)

pp.check(out.jags, actual = 'fit', new = 'fit.new')

#set.seed(13)
#m <- 10000
#chains <- samp[[1]][,1:4]
#chains2 <- samp[[2]][,1:4]

#postDraws <- chains[sample(nrow(chains),size=m,replace=TRUE),]



# Sample m draws of y.rep (n obs each) from likelihood p(y|theta), using thetas sampled above. Approximates posterior predictive distribution p(y.rep|y)
n <- nrow(full.ghcn)
y.rep <- matrix(NA, nrow=n, ncol=m)
for (i in 1:m){
  y.rep[,i] <- rnorm(n = n, postDraws[i,"Y[1]"], postDraws[i,"Y[3]"])
}

# check to see if the predicted yrep alighs with data y max
T1.y <- max(Y)
T1.yrep <- apply(y.rep, 2, max)
hist(T1.yrep)


abline(v=T1.y,col="red",lwd=2)


pppval.max <- sum(T1.yrep>=T1.y)/m
print(pppval.max)

# quick check to see how params compare to OLS 
summary(lm(Y ~ full.ghcn$DI.scaled))

# plot mcmc + the parameter distn
coda:::plot.mcmc(samp)
samp.basic.reg <- samp
# save a data frame with beta2 sensitivity and the 
samp.basic.reg.df <- summary(samp)
  
samp.basic.df.sum<- data.frame(mean = samp.df$statistics[2,"Mean"], 
           ci.low = samp.df$quantiles[2,"2.5%"], 
           ci.high = samp.df$quantiles[2,"97.5%"])
  


```

# Create random effects for sites:
```{r}
population_model_site_re <- "model{

# for each the overall population include re for sites:

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[i], inv.var) # where Yi is already log transformed

# function g()
gfunc[i] <- beta1[SITE[i]] + beta2[SITE[i]]*DI.scaled[i] + beta3[SITE[i]]*DBH.scaled[i]   # use Drought index as a scaled variable 

res[i] <- Y[i] - gfunc[i]   
emp.new[i] ~ dnorm(gfunc[i], sigma)
res.new[i] <- emp.new[i] - gfunc[i]
}



# Assume normal priors for betas, but generate a beta + alpha for each ageclass
for(s in 1:length(S)){
beta1[s] ~ dnorm(mu_beta1, inv_beta1)
beta2[s] ~ dnorm(mu_beta2, inv_beta2)
beta3[s] ~ dnorm(mu_beta3, inv_beta3)
}

# use normal hyperpriors for each hyperparamters 
mu_beta1 ~ dunif(-2, 2)
mu_beta2 ~ dunif(-2, 2)
mu_beta3 ~ dunif(-2, 2)

inv_beta1   ~ dgamma(0.0001, 0.0001)
sigma_beta1 <- 1/sqrt(inv_beta1)
inv_beta2   ~ dgamma(0.001, 0.001)
sigma_beta2 <- 1/sqrt(inv_beta2)
inv_beta3   ~ dgamma(0.001, 0.001)
sigma_beta3 <- 1/sqrt(inv_beta3)


# Non-informative Prior for the inverse population variances

#alpha_ref ~ dnorm(0,0.1)
inv.var   ~ dgamma(0.001, 0.001)
sigma     <- 1/sqrt(inv.var)


# Assess model fit using a sum-of-squares-type discrepancy
#for (i in 1:n) {
 #   predicted[i] <- gfunc[i]             # Predicted values
  #  residual[i] <- Y[i]-predicted[i]  # Residuals for observed data                                     
   # sq[i] <- pow(residual[i], 2)      # Squared residuals

# Generate replicate data and compute fit statistics for them
    #y.new[i]~dnorm(gfunc[i], inv.var)        # One new data set at each MCMC iteration
    #sq.new[i] <- pow(y.new[i]-predicted[i], 2)  # Squared residuals for new data
#}


#fit <- sum(sq[])              # Sum of squared residuals for actual data set
#fit.new <- sum(sq.new[])      # Sum of squared residuals for new data set
#test <- step(fit.new-fit) 		# Test whether new data set more extreme
#bpvalue <- mean(test) 		  	# Bayesian p-value

#Derived parameters
  fit <- sum(res[])
  fit.new <- sum(res.new[])


}"




#DI <- as.vector( full.ghcn$JJA.pdsi )
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x
full.ghcn$DBH.scaled = as.vector(scale(full.ghcn$DBH.x, center = TRUE, scale = TRUE))
site <- full.ghcn$site
SpecCode <- full.ghcn$SpecCode




reg.model.by_s <- jags.model(textConnection(population_model_site_re), 
                    data = list(Y=Y, n=n, DI.scaled = full.ghcn$DI.scaled, DBH.scaled = full.ghcn$DBH.scaled, SITE = as.numeric(full.ghcn$site), S = unique(full.ghcn$site)), n.chains = 3, n.adapt = 100)

update(reg.model.by_s, 1000); # Burnin for 1000 samples to start, then go higher later

samp <- coda.samples(reg.model.by_s, 
                     variable.names=c("beta1", "beta2","beta3","sigma", "sigma_beta1", "sigma_beta2", "sigma_beta3", "fit.new", "fit"), 
                    n.chains = 3, n.iter=20000)

summary(samp)
plot(samp)

gelman.diag(samp)

samp[1]

```


# model with cohort effects and re effects @ the cohort level:

```{r}
cohort_model_site_re <- "model{

# for each the overall population include re for sites:

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[i], inv.var) # where Yi is already log transformed

# function g()
gfunc[i] <- beta1[cohort[i]] + beta2[cohort[i]]*DI.scaled[i] + beta3[cohort[i]]*DBH.scaled[i]   # use Drought index as a scaled variable 
}



# Assume normal priors for betas, but generate a beta + alpha for each ageclass
for(s in 1:length(S)){
beta1[s] ~ dnorm(mu_beta1[indiv[s]], inv_beta1)
beta2[s] ~ dnorm(mu_beta2[indiv[s]], inv_beta2)
beta3[s] ~ dnorm(mu_beta3[indiv[s]], inv_beta3)
}


for(j in 1:length(ind)){
# use normal hyperpriors for each hyperparamters 
mu_beta1[j] ~ dnorm(0, 0.5)
mu_beta2[j] ~ dnorm(0, 0.5)
mu_beta3[j] ~ dnorm(0, 0.5)

}

inv_beta1   ~ dgamma(0.001, 0.001)
sigma_beta1 <- 1/sqrt(inv_beta1)
inv_beta2   ~ dgamma(0.001, 0.001)
sigma_beta2 <- 1/sqrt(inv_beta2)
inv_beta3   ~ dgamma(0.001, 0.001)
sigma_beta3 <- 1/sqrt(inv_beta3)


# Non-informative Prior for the inverse population variances

#alpha_ref ~ dnorm(0,0.1)
inv.var   ~ dgamma(0.0001, 0.0001)
sigma     <- 1/sqrt(inv.var)

}"




#DI <- as.vector( full.ghcn$JJA.pdsi )
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x
full.ghcn$DBH.scaled = as.vector(scale(full.ghcn$DBH.x, center = TRUE, scale = TRUE))
site <- full.ghcn$site
SpecCode <- full.ghcn$SpecCode




cohort.model.re <- jags.model(textConnection(cohort_model_site_re), 
                    data = list(Y=Y, n=n, DI.scaled = full.ghcn$DI.scaled, DBH.scaled = full.ghcn$DBH.scaled, cohort = as.numeric(full.ghcn$ageclass), S = unique(full.ghcn$ageclass),
                    indiv = as.numeric(full.ghcn$ID), ind = unique(full.ghcn$ID)), n.chains = 3, n.adapt = 100)

update(cohort.model.re, 20000); # Burnin for 1000 samples to start, then go higher later

samp <- coda.samples(cohort.model.re, 
                     variable.names=c("beta1", "beta2","beta3","sigma", "sigma_beta1", "sigma_beta2", "sigma_beta3", "mu_beta1", "mu_beta2", "mu_beta3"), 
                    n.chains = 3, n.iter=10000, thin = 10)

summary(samp)
traceplot(samp)
gelman.diag(samp)


```

## Bayesian heirarchical liner regression where alpha and beta coeffiencents are specific to savanna and forest cover:
# this model explores whether stand structure (savanna or forest) has an impact on the slope and intercept response to climate

```{r}


Y <- as.vector(log(full.ghcn$RWI)) 
# create table of savanna and forest

for.df <- data.frame(site = unique(full.ghcn$site), 
           fortype = c("forest", "savanna", "forest" ,"forest", "savanna", "forest", "savanna", "savanna", "forest", "savanna", "savanna", "savanna", "savanna", "savanna", "forest", "savanna"))

full.ghcn <- merge(full.ghcn, for.df, by = "site")
full.ghcn$fornum <- ifelse(full.ghcn$fortype %in% "forest", 1, 2)
# standardise predictor variables to have mean 0 and sd = 1
#DI.scaled = scale(full.ghcn$JJA.pdsi, center= TRUE, scale=TRUE)
#T.scaled = scale(full.ghcn$JUNTmax, center= TRUE, scale=TRUE)
#full.ghcn$DI.scaled = as.vector(scale(full.ghcn$JJA.pdsi, center = TRUE, scale = TRUE))
ggplot(full.ghcn, aes(DI.scaled,log(RWI), color = ageclass))+geom_point(size = 0.2)+geom_smooth(method = "lm")+facet_wrap(~fortype)


png(height= 5, width = 7, units = "in", res = 300, "outputs/two_age_class_logbai_past_mod.png")
ggplot(full.ghcn[full.ghcn$RWI <= 6200, ], aes(JJA.pdsi, log(RWI), color = ageclass))+geom_point(size = 0.02)+geom_smooth(method = "lm")+facet_wrap(~fortype)
dev.off()

#DI <- as.vector( full.ghcn$JJA.pdsi )
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x
full.ghcn$DBH.scaled = as.vector(scale(full.ghcn$DBH.x, center = TRUE, scale = TRUE))
site <- full.ghcn$site
SpecCode <- full.ghcn$SpecCode

summary(lm(Y ~ log(DBH)))

summary(lm(Y ~ log(DBH) + log((pdsi + 10)/20) + log(JunTmax)  ))
summary(lmer(Y ~ DI.scaled+ (DI.scaled|ageclass)))
# population model for the response of each BAI to each year of climate:
population_model_age_class <- "model{

# for each type of stand structure:



# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[fortype[i]], inv.var) # where Yi is already log transformed

# function g()
gfunc[i] <- alpha[fortype[i]]+ beta2[fortype[i]] *DI.scaled[i]    # use Drought index as a scaled variable 
}



# Assume normal priors for betas, but generate a beta + alpha for each ageclass
for(s in 1:length(S)){
alpha[s] ~ dnorm(mu_alpha, inv_alpha)
beta2[s] ~ dnorm(mu_beta, inv_beta)
}

# use normal hyperpriors for each hyperparamters 
mu_alpha ~ dunif(-2, 2)
mu_beta ~ dunif(-2, 2)

inv_alpha   ~ dgamma(0.0001, 0.0001)
sigma_alpha <- 1/sqrt(inv_alpha)
inv_beta   ~ dgamma(0.001, 0.001)
sigma_beta <- 1/sqrt(inv_beta)



# Non-informative Prior for the inverse population variances

#alpha_ref ~ dnorm(0,0.1)
inv.var   ~ dgamma(0.001, 0.001)
sigma     <- 1/sqrt(inv.var)

}"

lmer.age <- lmer(Y ~ DI.scaled + (1|ageclass == 1))
ranef(lmer.age)
ggplot(full.ghcn[full.ghcn$RWI < 5500, ], aes(DBH.x, RWI , color = ageclass))+geom_point(size = 0.02)+stat_smooth(method = "lm")
indiv <- lmer(log(RWI) ~ DI.scaled + (DI.scaled|ID), data = full.ghcn)

initsList

# now 
reg.model.by_s <- jags.model(textConnection(population_model_age_class), 
                    data = list(Y=Y, n=n, DI.scaled = full.ghcn$DI.scaled, DBH.scaled = full.ghcn$DBH.scaled, fortype = ageclass, S = unique(ageclass)), n.chains = 3, n.adapt = 100)

update(reg.model.by_s, 40000); # Burnin for 1000 samples to start, then go higher later

samp <- coda.samples(reg.model.by_s, 
                     variable.names=c("beta", "beta2","alpha","sigma", "sigma_alpha"), 
                    n.chains = 3, n.iter=20000, thin = 10)

summary(samp)
gelman.diag(samp)
traceplot(samp)
plot(samp)

# quick check to see how params compare to OLS 
#summary(lm(Y ~ full.ghcn$DI.scaled))

# plot mcmc + the parameter distn
plot(samp)
gelman.diag(samp)
acfplot(samp, aspect = 2)


gelman.plot(samp)
# evaluate MCMC convergence:


# save a data frame with beta2 sensitivity and the 
samp.basic.reg.df <- summary(samp)
  
samp.basic.df.sum<- data.frame(mean = samp.df$statistics[2,"Mean"], 
           ci.low = samp.df$quantiles[2,"2.5%"], 
           ci.high = samp.df$quantiles[2,"97.5%"])
  



# lets do the model where stand only affects the slope and not the intercept, perhaps this will help it converge:

population_model_stand <- "model{

# for each type of stand structure:



# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[fortype[i]], inv.var) # where Yi is already log transformed


# function g()
gfunc[i] <- alpha +  beta[fortype[i]]*DI.scaled[i] # use Drought index as a scaled variable 
}

# Assume normal priors for betas, but generate a beta + alpha for each forest type s
for(s in 1:length(S)){

beta[s] ~ dnorm(mu_beta, inv_beta)
}

# use normal hyperpriors for each hyperparamters 
#mu_alpha ~ dnorm(0, 0.1)
mu_beta ~ dnorm(0, 0.1)

#inv_alpha   ~ dgamma(0.001, 0.001)
#sigma_alpha <- 1/sqrt(inv_alpha)
inv_beta   ~ dgamma(0.001, 0.001)
sigma_beta <- 1/sqrt(inv_beta)



# Non-informative Prior for the inverse population variances
inv.var   ~ dgamma(0.001, 0.001)
sigma     <- 1/sqrt(inv.var)
alpha ~ dnorm(0, 0.1)

}"

# now 
reg.model.by_s <- jags.model(textConnection(population_model_stand), 
                    data = list(Y=Y, n=n, DI.scaled = full.ghcn$DI.scaled, fortype =    full.ghcn$fornum, S = unique(full.ghcn$fornum)))

update(reg.model.by_s, 1000); # Burnin for 1000 samples to start, then go higher later

samp <- coda.samples(reg.model.by_s, 
                     variable.names=c("beta","alpha", "sigma"), 
                    n.chains = 4, n.iter=20000)

summary(samp)



```
# this model assumes all trees across all species + sites have the same relationship with summer PDSI + JUNTmax

```{r}
summary(full.ghcn)

full.ghcn <- full.ghcn[!is.na(full.ghcn$RWI) & full.ghcn$DBH.x > 0 & !is.na(full.ghcn$DBH.x),]
Y <- as.vector(log(full.ghcn$RWI)) 

pdsi <- as.vector( full.ghcn$JJA.pdsi)
JunTmax <- as.vector( full.ghcn$JUNTmax)
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x

model_string_pdsi_tmax <- "model{

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[i], inv.var)

# function g()
gfunc[i] <- beta[1] +  beta[2]*pdsi[i] + beta[3]*JunTmax[i]
}

# Assume normal priors for betas (because they could be negative or positive)
for(j in 1:3){
beta[j] ~ dnorm(0,0.001)
}



# Prior for the inverse variances
inv.var   ~ dgamma(0.1, 0.1)
sigma     <- 1/sqrt(inv.var)



}"

reg.model.Drought.Tmax <- jags.model(textConnection(model_string_pdsi_tmax), 
                    data = list(Y=Y, n=n, pdsi = pdsi, JunTmax = JunTmax))

update(reg.model.Drought.Tmax, 10000, progress.bar="none"); # Burnin for 10000 samples

samp <- coda.samples(reg.model.Drought.Tmax, 
                     variable.names=c("beta","sigma"), 
                    n.chains = 3, n.iter=20000, progress.bar="none")

summary(samp)


# plot mcmc + the parameter distn
plot(samp)

# save a data frame with beta2 sensitivity and the 
samp.df <- summary(samp)
  
samps.df.sum<- data.frame(mean = samp.df$statistics[2,"Mean"], 
           ci.low = samp.df$quantiles[2,"2.5%"], 
           ci.high = samp.df$quantiles[2,"97.5%"])
  

ggplot(samps.df.sum, aes("sens",mean))+geom_bar(stat = "identity")+
  geom_errorbar(aes(ymin=ci.low, ymax = ci.high), size = 0.2, width = 0.2)

```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


# this model assumes all trees across all species + sites have the same relationship with summer PDSI + JUNTmax + DBH, but has a power function as the process model:



```{r}
summary(full.ghcn)

full.ghcn <- full.ghcn[!is.na(full.ghcn$RWI) & full.ghcn$DBH.x > 0 & !is.na(full.ghcn$DBH.x),]


Y <- as.vector(log(full.ghcn$RWI)) 

pdsi <- as.vector( full.ghcn$JJA.pdsi)
JunTmax <- as.vector( full.ghcn$JUNTmax)
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x

model_string_DBH_power <- "model{

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[i], inv.var)

# function g()
gfunc[i] <- beta[1] +  beta[1]*log(DBH[i])
}

# Assume normal priors for betas (because they could be negative or positive)
#for(j in 1:2){
beta[1] ~ dnorm(0, 0.002)
beta[2] ~ dnormal(0, 0.002)
#}



# Prior for the inverse variances
inv.var   ~ dgamma(0.01, 0.01)
sigma     <- 1/sqrt(inv.var)


}"

reg.model.dbh.power <- jags.model(textConnection(model_string_DBH_power), 
                    data = list(Y=Y, n=n, DBH = DBH))

update(reg.model.dbh.power, 1000); # Burnin for 1000 samples

samp <- coda.samples(reg.model.dbh.power, 
                     variable.names=c("beta","sigma"), 
                    n.chains = 3, n.iter=2000)

summary(samp)


# plot mcmc + the parameter distn
plot(samp)

# save a data frame with beta2 sensitivity and the 
samp.df <- summary(samp)
  
samp.dbh.pwr <- do.call(rbind.data.frame, samp)


# compare to the estimates from OLS:

summary(lm(Y ~ log(DBH)))
```
# this model assumes all trees across all species + sites have the same relationship with summer PDSI + JUNTmax + DBH, but has a power function as the process model:

```{r}
summary(full.ghcn)

full.ghcn <- full.ghcn[!is.na(full.ghcn$RWI) & full.ghcn$DBH.x > 0 & !is.na(full.ghcn$DBH.x),]
Y <- as.vector((full.ghcn$RWI)) 

pdsi <- as.vector( full.ghcn$JJA.pdsi)
JunTmax <- as.vector( full.ghcn$JUNTmax)
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x

model_string_pdsi_tmax_DBH_power <- "model{

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dlnorm(gfunc[i], inv.var)

# function g()
gfunc[i] <- beta[1] +  beta[2]*log((pdsi[i]+10)/20) + beta[3]*log(JunTmax[i]) + beta[4]*log(DBH[i])
}

# Assume normal priors for betas (because they could be negative or positive)
for(j in 1:4){
beta[j] ~ dlnorm(0,0.001)
}



# Prior for the inverse variances
inv.var   ~ dgamma(0.1, 0.1)
sigma     <- 1/sqrt(inv.var)


}"

reg.model.Drought.Tmax.dbh.power <- jags.model(textConnection(model_string_pdsi_tmax_DBH_power), 
                    data = list(Y=Y, n=n, pdsi = pdsi, JunTmax = JunTmax, DBH = DBH))

update(reg.model.Drought.Tmax.dbh.powe, 10000, progress.bar="none"); # Burnin for 10000 samples

samp <- coda.samples(reg.model.Drought.Tmax.dbh, 
                     variable.names=c("beta","sigma"), 
                    n.chains = 3, n.iter=20000, progress.bar="none")

summary(samp)


# plot mcmc + the parameter distn
plot(samp)

# save a data frame with beta2 sensitivity and the 
samp.df <- summary(samp)
  

```
# model assumes a power function relationship between growth ~ beta0x(DBH^beta1)x(DBH^2^beta2)x(droghtindex^beta3)x(JunTmax^beta4)
# will need to scale pdsi such that: drought index == ((pdsi+10)/20)

```{r}
summary(full.ghcn)

full.ghcn <- full.ghcn[!is.na(full.ghcn$RWI) & full.ghcn$DBH.x > 0 & !is.na(full.ghcn$DBH.x),]
Y <- as.vector(log(full.ghcn$RWI)) 

pdsi <- as.vector( (full.ghcn$JJA.pdsi + 10)/20)
JunTmax <- as.vector( full.ghcn$JUNTmax)
ageclass <- as.numeric( full.ghcn$ageclass)
Age <- as.numeric( full.ghcn$Age)
n     <- length(full.ghcn$RWI)
DBH <- full.ghcn$DBH.x

model_string_pdsi_tmax_DBH <- "model{

# Likelihood
for(i in 1:n){
# process model
Y[i]   ~ dnorm(gfunc[i], inv.var)

# function g()
gfunc[i] <- beta[1] +  beta[2]*pdsi[i] + beta[3]*JunTmax[i] + beta[4]*DBH[i]
}

# Assume normal priors for betas (because they could be negative or positive)
for(j in 1:4){
beta[j] ~ dnorm(0,0.001)
}



# Prior for the inverse variances
inv.var   ~ dgamma(0.1, 0.1)
sigma     <- 1/sqrt(inv.var)



}"

reg.model.Drought.Tmax.dbh <- jags.model(textConnection(model_string_pdsi_tmax_DBH), 
                    data = list(Y=Y, n=n, pdsi = pdsi, JunTmax = JunTmax, DBH = DBH))

update(reg.model.Drought.Tmax.dbh, 10000, progress.bar="none"); # Burnin for 10000 samples

samp <- coda.samples(reg.model.Drought.Tmax.dbh, 
                     variable.names=c("beta","sigma"), 
                    n.chains = 3, n.iter=20000, progress.bar="none")

summary(samp)


# plot mcmc + the parameter distn
plot(samp)

# save a data frame with beta2 sensitivity and the 
samp.df <- summary(samp)
  

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
