---
title: "KF_tree_rings"
author: "Kelly Heilman"
date: "February 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Kalman Filter/Time varying Regression model to estimate tree ring regressions 
code  based off an example from:
<http://lalas.github.io/quantitativeThoughts/r/2014/09/01/dlmTutorial.html#topic2.1>

Here we read in example data
```{r cars}
library(dplR)
Bonanza <- read.tucson("/Users/kah/Documents/TreeRings/cofecha/BONww.rwl", header = TRUE)
BON.rwi <- detrend(rwl = Bonanza, method = "Spline")

#create chronology of sites
BON.crn <- chron(BON.rwi, prefix = paste('BON'))
plot(BON.crn)

head(BON.crn)

# read in molten df with climate
molten.HIC<- read.csv("/Users/kah/Documents/TreeRings/data/molten.hic.csv")

```



```{r pressure, echo=FALSE}
library(PerformanceAnalytics, quietly = TRUE,  warn.conflicts = FALSE)
library(dlm)

det.full.df <- read.csv("/Users/kah/Documents/TreeRings/outputs/data/full_det_ghcn_rwi.csv")
val <- det.full.df[det.full.df$site %in% "BON",]
clim <- "Jul.pdsi"
site.code<- "BON"
kalman.plot <- function(val, clim, site.code){

  
  
    # convert to xts objects
    tr <- as.xts(val[,c("RWI")], order.by = as.yearqtr(val$year))
    
    pdsi <- as.xts(val[,c( clim )], order.by = as.yearqtr(val$year))
    
    #tmin <- as.xts(val[,c( clim2 )], order.by = as.yearqtr(val$Year))
    
    
    # Specifying a set model parameters
    s2_obs = 1      # Variance of observations
    s2_alpha = 0.0001 # Variance of the alpha regression parameter
    s2_beta = 0.0001  # Variance of the beta regression parameter
    s2_beta2 = 0.0001 
  
   
    tvp.dlm = dlmModReg(X=pdsi, addInt=TRUE, dV=s2_obs, dW=c(s2_alpha, s2_beta))
    
    #Now that we have defined a model, we can view its different component:
    
    # looking at the various component
    tvp.dlm[c("FF","V","GG","W","m0","C0")]
    tvp.dlm[c("JFF","JV","JGG","JW")]
    
    #If we were to do a simple linear regression (Ordinary Least Square fit - constant equity beta), we would do something like
    
    
    ols.fit <- lm(tr ~ pdsi )
    summary(ols.fit)
    
    start.vals = c(0,0,0)
    # Names ln variance of: observation y, alpha and beta (corresponding intercept and slope of y (tr index) with respect to X (PDSI))
    names(start.vals) = c("lns2_obs", "lns2_alpha", "lns2_beta")
    
    # function to build Time Varying Parameter state space model
    buildTVP <- function(parm, x.mat){
        parm <- exp(parm)
      return( dlmModReg(X=x.mat, dV=parm[1], dW=c(parm[2], parm[3])) )
    }
    
    # Estimate the model
    TVP.mle = dlmMLE(y=tr, parm=start.vals, x.mat=pdsi, build=buildTVP, hessian=T)
    
    # get sd estimates
    se2 <- sqrt(exp(TVP.mle$par))
    names(se2) = c("s_obs", "s_alpha", "s_beta")
    sqrt(se2)
    
    
    # Build fitted ss model, passing to it sp500 as the matrix X in the model
    TVP.dlm <- buildTVP(TVP.mle$par, pdsi)
    #Filtering and Smooting:
    #Filtering Optimal estimates of θtθt given information available at time tt, It={y1,...,yt}It={y1,...,yt}
    #Smoothing Optimal estimates of θtθt given information available at time TT, IT={y,...,yT}IT={y,...,yT}
    #Now that we have obtained model estimates, and build the optimal model, we can filter the data through it, to obtain filtered values of the state vectors, together with their variance/co-variance matrices.
    
    TVP.f <- dlmFilter(y = tr, mod = TVP.dlm)
    class(TVP.f)
    
    names(TVP.f)
    
    #Similarly, to obtained the smoothed values of the state vectors, together with their variance/co-variance matrices; using knowledge of the entire series
    
    # Optimal estimates of θ_t given information available at time T.
    TVP.s <- dlmSmooth(TVP.f)
    #class(TVP.s)
    
    
    
    #Plotting the results (smoothed values)
    #Now that we have obtained the smoothed values of the state vectors, we can draw them as:
    
    
    
    # extract smoothed states - intercept and slope coefs
    alpha.s = xts(TVP.s$s[-1,1,drop=FALSE], as.yearqtr( rownames(TVP.s$s[-1,])))
    
    beta.s  = xts(TVP.s$s[-1,2,drop=FALSE], as.yearqtr(rownames(TVP.s$s[-1,])))
    colnames(alpha.s) = "alpha"
    colnames(beta.s)  = "beta"
    #Extracting the std errors and constructing the confidence band
    
    # extract std errors - dlmSvd2var gives list of MSE matrices
    mse.list = dlmSvd2var(TVP.s$U.S, TVP.s$D.S)
    se.mat = t(sapply(mse.list, FUN=function(x) sqrt(diag(x))))
    se.xts = xts(se.mat[-1, ], index(beta.s))
    colnames(se.xts) = c("alpha", "beta")
    a.u = alpha.s + 1.96*se.xts[, "alpha"]
    a.l = alpha.s - 1.96*se.xts[, "alpha"]
    b.u = beta.s  + 1.96*se.xts[, "beta"]
    b.l = beta.s  - 1.96*se.xts[, "beta"]
    #And plotting the results with +/- 2 times the standard deviation
    par(mfrow = c(2,1))
    
    # plot smoothed estimates with +/- 2*SE bands
    chart.TimeSeries(cbind(alpha.s, a.l, a.u), main=paste("Smoothed estimates of alpha for", site.code), ylim=c(0,2),
                     colorset=c(1,2,2), lty=c(1,2,2),ylab=expression(alpha),xlab="")
    
    chart.TimeSeries(cbind(beta.s, b.l, b.u), main=paste("Smoothed estimates of beta for", site.code),
                     colorset=c(1,2,2), lty=c(1,2,2),ylab=expression(beta),xlab="")

    
alpha.df <- data.frame(dateTime = index(se.xts), alpha = alpha.s, upr = a.u, lwr = a.l)
names(alpha.df) <- c("dateTime", "alpha", "upr", "lwr")
beta.df  <- data.frame(dateTime = index(se.xts), beta = beta.s, upr = b.u, lwr = b.l)
names(beta.df) <- c("dateTime", "beta", "upr", "lwr")

ggplot(data = alpha.df, aes(dateTime, alpha) ) + geom_point () + geom_line() + geom_ribbon(data=alpha.df, aes(ymin=lwr,ymax=upr), alpha=0.3) + labs(x = "year", y = expression(alpha), title = expression(paste("State Space Values of ", alpha, " over Time")))

ggplot(data = beta.df, aes(dateTime, beta) ) + geom_point (data = beta.df, aes(dateTime, beta) ) + geom_line() + geom_ribbon(data=beta.df , aes(ymin=lwr,ymax=upr), alpha=0.3) + labs(x = "year", y = expression(beta), title = expression(paste("State Space Values of ", beta, " over Time")))

}



#using molten outputs form climate_growth_reg_chron.R

#X11(width = 12)
pdf("/Users/kah/Documents/TreeRings/outputs/Kalman_regressions.pdf")
kalman.plot(val = molten.UNC, clim = "Jul.pdsi", site.code = "UNC")
kalman.plot(val = molten.BON, clim = "Jul.pdsi", site.code = "BON")
kalman.plot(val = val, clim = "Jul.pdsi", site.code = "HIC")
kalman.plot(val = molten.STC, clim = "Jul.pdsi", site.code = "STC")
kalman.plot(val = molten.TOW, clim = "Jul.pdsi", site.code = "TOW")
kalman.plot(val = molten.COR, clim = "Jul.pdsi", site.code = "COR")
kalman.plot(val = molten.ENG, clim = "Jul.pdsi", site.code = "ENG")
kalman.plot(val = molten.COR, clim = "Jul.pdsi", site.code = "COR")
dev.off()


kalman.plot(val = molten.UNC, clim = "PDSI", site.code = "UNC")

kalman.plot(val = molten.BON, clim = "PDSI", site.code = "BON")

kalman.plot(val = molten.HIC, clim = "PDSI", site.code = "HIC")
kalman.plot(val = molten.STC, clim = "PDSI", site.code = "STC")
kalman.plot(val = molten.TOW, clim = "PDSI", site.code = "TOW")
kalman.plot(val = molten.COR, clim = "PDSI", site.code = "COR")
kalman.plot(val = molten.ENG, clim = "PDSI", site.code = "ENG")
kalman.plot(val = molten.COR, clim = "PDSI", site.code = "COR")


```

#####################
jags model of tree ring growth 
```{r}
TreeRingModel = "
model{

  #### Data Model -- linear regression of rwi with drought
 for(i in 1:N){
 Y[i] ~ dnorm(mu[i], tau) # Observation
mu[i] <- alpha + beta * (x[i] - mean(x[])) # System Process Model
 }

 #priors
  # Distributions are set to all parameters
 sigma <- 1/sqrt(tau)
  alpha ~ dnorm(0, 1.0E-6)
  beta ~ dnorm(0, 1.0E-6)
  tau ~ dgamma(1.0E-3, 1.0E-3)
}
"

```
Define data and priors
```{r}
data <- list(Y=molten.HIC$value, x = molten.HIC$Jul.pdsi, N=length(Y))
```

Define intitial conditions
```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  Y.samp = sample(Y,length(Y),replace=TRUE)
  init[[i]] <- list(tau=1/var(diff(Y.samp)),beta=5/var(Y.samp), alpha = 5/var(Y.samp))
}
```

Jags model
```{r}
j.model   <- jags.model (file = textConnection(TreeRingModel),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

assess convergence
```{r}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("alpha","beta"),
                                n.iter = 1000)
plot(jags.out)
```


```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","alpha","beta"),
                                n.iter = 10000)
```

Visualize posteriors 

```{r}
ciEnvelope <- function(x,ylo,yhi,...){
  polygon(cbind(c(x, rev(x), x[1]), c(ylo, rev(yhi),
                                      ylo[1])), border = NA,...) 
}
out <- as.matrix(jags.out)
ci <- apply(exp(out[,3:ncol(out)]),2,quantile,c(0.025,0.5,0.975))

time <- molten.HIC$Year
plot(time,ci[2,],type='n',ylim=range(y,na.rm=TRUE),ylab="RWI",log='y')
ciEnvelope(time,ci[1,],ci[3,],col="lightBlue")
points(time,y,pch="+",cex=0.5)

```

posterieors
```{r}
layout(matrix(c(1,2,3,3),2,2,byrow=TRUE))
hist(1/sqrt(out[,1]),main=colnames(out)[1])
hist(1/sqrt(out[,2]),main=colnames(out)[2])
plot(out[,1],out[,2],pch=".",xlab=colnames(out)[1],ylab=colnames(out)[2])
cor(out[,1:2])
```

#######################################
# defining KF by hand--still needs work
# define the kalman filter here
```{r}
KalmanFilter <- function(M,mu0,P0,Q,R,Y){
  ##Inputs:
  ##  M   = model matrix
  ##  mu0 = initial condition mean vector
  ##  P0  = initial condition covariance matrix
  ##  Q   = process error covariance matrix
  ##  R   = observation error covariance matrix
  ##  Y   = observation matrix (with missing values as NAs), time as col's
  
  ## Output: (as a list)
  ##  mu.f, mu.a  = state mean vector for (a)nalysis and (f)orecast steps
  ##  P.f, P.a    = state covariance matrix for a and f
  
  ## storage
  nstates = nrow(Y)  
  nt = ncol(Y)
  mu.f  = matrix(NA,nstates,nt+1)  ## forecast mean for time t
  mu.a  = matrix(NA,nstates,nt)  ## analysis mean for time t
  P.f  = array(NA,c(nstates,nstates,nt+1))  ## forecast variance for time t
  P.a  = array(NA,c(nstates,nstates,nt))  ## analysis variance for time t

  ## initialization
  mu.f[,1] = mu0
  P.f[,,1] = P0
  I = diag(1,nstates)

  ## run updates sequentially for each observation.
  for(t in 1:nt){

    ## Analysis step: combine previous forecast with observed data
    obs = !is.na(Y[,t]) ## which Y's were observed?
    if(any(obs)){
      H <- I[obs,]                                                        ## observation matrix
      K <- P.f[,,t] %*% t(H) %*% solve(H%*%P.f[,,t]%*%t(H) + R[obs,obs])  ## calculate Kalman gain
      mu.a[,t] <- mu.f[,t] + K%*%(Y[obs,t] - H %*% mu.f[,t])              ## update mean
      P.a[,,t] <- (1-K %*% H)*P.f[,,t]                                    ## update covariance
    } else {
      ##if there's no data, the posterior is the prior
      mu.a[,t] = mu.f[,t]
      P.a[,,t] = P.a[,,t]
    }

    ## Forecast step: predict to next step from current
    mu.f[,t+1] = M%*%mu.a[,t]
    P.f[,,t+1] = Q + M*P.a[,,t]*t(M)
  
  }
  
  return(list(mu.f=mu.f,mu.a=mu.a,P.f=P.f,P.a=P.a))
}

ciEnvelope <- function(x,ylo,yhi,...){
  polygon(cbind(c(x, rev(x), x[1]), c(ylo, rev(yhi),
                                      ylo[1])), border = NA,...) 
}
```

```{r}
time = molten.HIC$Year
states = c("value")
nstates = length(states)
y = t(molten.HIC[,states])

## define adjacency between states slected
adj = matrix(c(0),nstates,nstates,byrow=TRUE)

## plot time-series from states
plot(time,1:length(time),type='n',ylab="Tree Ring Index Index",lwd=2,log='y',ylim=range(y,na.rm=TRUE))
for(i in 1:nstates){
  lines(time,y[i,],col=i,lwd=2)
}
legend("topleft",legend=states,lwd=2,col=1:nstates)

## log transform data
Y   = log10(y)

## load parameters (assume known)
load("data/KFalpha.params.Rdata")

## options for process model 
alpha = 0       ## assume no spatial flux
#alpha = 0.05    ## assume a large spatial flux
M = adj*alpha + diag(1-alpha*apply(adj,1,sum))  ## random walk with flux

## options for process error covariance
Q = tau_proc            ## full covariance matrix
Q = diag(diag(Q))       ## diagonal covariance matrix
Q = 0.0131
## observation error covariance (assumed independent)  
R = diag(tau_obs,nstates) 

## prior on first step, initialize with long-term mean and covariance
mu0 = apply(Y,1,mean,na.rm=TRUE)
P0 = cov(t(Y),use="pairwise.complete.obs")

## Run Kalman Filter
KF00 = KalmanFilter(M,mu0,P0,Q,R,Y)
```

```{r}
attach(KF00)
nt = length(time)

### plot ANALYSIS mean & CI time-series
par(mfrow=c(3,1))
for(i in 1){
  ci = rbind(mu.a[i,]-1.96*sqrt(P.a[i,i,]),mu.a[i,]+1.96*sqrt(P.a[i,i,]))
  plot(time,mu.a[i,],ylim=range(ci,na.rm=TRUE),type='n',main=states[i])
  ciEnvelope(time,ci[1,],ci[2,],col="lightBlue")
  lines(time,mu.a[i,],col=4)
  lines(time,Y[i,])
}

## plot ANALYSIS and FORECAST variance time-series
par(mfrow=c(3,1))
for(i in 1){
  plot(time,sqrt(P.a[i,i,]),ylim=c(0,sqrt(max(c(P.a[i,i,],P.f[i,i,])))),main=states[i],xlab="Time",
       ylab="Std Error",type='l')
  lines(time,sqrt(P.f[i,i,1:nt]),col=2)
  points(time[is.na(Y[i,])],rep(0,nt)[is.na(Y[i,])],pch="*",col=3) ## flag's the zero's
  legend("topright",legend=c("Analysis","Forecast","NAs"),col=1:3,lty=c(1,1,NA),pch=c(NA,NA,1),cex=1.4)
}
```

## compare to moving window correlations:
```{r}
library(treeclim)

# we will us the dcc function in the tree clim package, but this funtion takes monthly data:

#read climate
IL.clim <- read.csv("data/NE_illinois_climdiv.csv") #Hickory Grove, Sandwich, Glacial park
WIse.clim <- read.csv("data/south_east_wi_climdiv.csv") #pleasant prairie 
MNwc.clim <- read.csv("data/West_central_MN_nclimdiv.csv") #Bonanza praire, Duboix
WIsc.clim <- read.csv("data/south_central_WI_climdiv.csv") #pleasant valley conservancy
MNec.clim <- read.csv("data/East_Central_MN_CDODiv5039587215503.csv") #townsend woods 
MNec.clim <- read.csv("data/East_Central_MN_CDODiv5039587215503.csv") #St. Croix savanna
MNse.clim <- read.csv("data/South_East_MN_CDO.csv") # for mound prairie

# We need the raw data:
Glacial <- read.tucson("C:/Users/JMac/Documents/Kelly/crossdating/data/cofecha/GLA.rwl")
Hickory <- read.tucson ("./cofecha/HICww.rwl")
Bonanza <- read.tucson("./cofecha/BONww.rwl")
Pleasant <- read.tucson('./cofecha/PLEww.rwl')
Townsend <- read.tucson('./cofecha/tow/TOWww.rwl')
StCroix <- read.tucson('./cofecha/STCww.rwl')
Coral <- read.tucson('C:/Users/JMac/Documents/Kelly/crossdating/data/cofecha/COR.rwl')
Uncas <- read.tucson("C:/Users/JMac/Documents/Kelly/crossdating/data/cofecha/UNC.rwl")
Englund <- read.tucson("C:/Users/JMac/Documents/Kelly/crossdating/data/cofecha/ENG.rwl")
Mound <- read.tucson("C:/Users/JMac/Documents/Kelly/crossdating/data/cofecha/MOU.rwl")




source("R/read_detrend_rwl.R")

Glacial <- read_detrend_rwl(Glacial, "Glacial")
Hickory <- read_detrend_rwl(Hickory, "Hickory")
Bonanza <- read_detrend_rwl(Bonanza, "Bonanza")
Pleasant <- read_detrend_rwl(Pleasant, "Pleasant")
Townsend <- read_detrend_rwl(Townsend, "Townsend")
StCroix <- read_detrend_rwl(StCroix, "StCroix")
Coral <- read_detrend_rwl(Coral, "Coral")
Uncas <- read_detrend_rwl(Uncas, "Uncas")
Englund <- read_detrend_rwl(Englund, "Englund")
Mound <- read_detrend_rwl(Mound, "Mound")

# moving correlations between climate and tree growth:

a<- dcc(Hickory, IL.clim[1:1452,c("Year", "Month", "PCP")], dynamic = 'moving', win_size = 35, win_offset = 5)
a

plot(a)
traceplot(a)


#these funcitons print out plots time moving correlations for all of the climate parameters

clim.cor<- function(climate, chron, site.name){

PREC <- climate[,c('Year', 'Month', 'PCP')]
PREC$PCP <- PREC$PCP*25.54
PREC <- PREC[1:1452,]


hic.pdsi.static <- dcc(chron, PREC, dynamic = 'static', win_size = 35, win_offset = 5)
pdf(paste0('outputs/correlations/moving_site_cors/PREC_', site.name,'dynamic.pdf'))
print(plot(hic.pdsi.static))
#g_test(hic.pdsi.moving)
#traceplot(hic.pdsi.moving)
#plot(skills(hic.pdsi.moving))


hic.pdsi.moving <- dcc(chron, PREC, dynamic = 'moving', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.moving))
#g_test(hic.pdsi.moving)
print(traceplot(hic.pdsi.moving))
#plot(skills(hic.pdsi.moving))
dev.off()

#PDSI
PDSI <- climate[,c('Year', 'Month', 'PDSI')]

PDSI <- PDSI[1:1452,]

pdf(paste0('outputs/correlations/moving_site_cors/PDSI_', site.name,'dynamic.pdf'))
hic.pdsi.static <- dcc(chron, PDSI, dynamic = 'static', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.static))
#g_test(hic.pdsi.moving)
#traceplot(hic.pdsi.moving)
#plot(skills(hic.pdsi.moving))

hic.pdsi.moving <- dcc(chron, PDSI, dynamic = 'moving', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.moving))
#g_test(hic.pdsi.moving)
print(traceplot(hic.pdsi.moving))
#plot(skills(hic.pdsi.moving))
dev.off()

#TAVG
TAVG <- climate[,c('Year', 'Month', 'TAVG')]

TAVG <- TAVG[1:1452,]

pdf(paste0('outputs/correlations/moving_site_cors/TAVG_', site.name,'dynamic.pdf'))
hic.pdsi.static <- dcc(chron, TAVG, dynamic = 'static', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.static))


hic.pdsi.moving <- dcc(chron, TAVG, dynamic = 'moving', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.moving))
#g_test(hic.pdsi.moving)
print(traceplot(hic.pdsi.moving))
dev.off()


#TMAX
TMAX <- climate[,c('Year', 'Month', 'TMAX')]

TMAX <- TMAX[1:1452,]

pdf(paste0('outputs/correlations/moving_site_cors/TMAX_', site.name,'dynamic.pdf'))
hic.pdsi.static <- dcc(chron, TMAX, dynamic = 'static', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.static))


hic.pdsi.moving <- dcc(chron, TMAX, dynamic = 'moving', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.moving))
#g_test(hic.pdsi.moving)
print(traceplot(hic.pdsi.moving))
#plot(skills(hic.pdsi.moving))
dev.off()

#TMIN
TMIN <- climate[,c('Year', 'Month', 'TMIN')]

TMIN <- TMIN[1:1452,]

pdf(paste0('outputs/correlations/moving_site_cors/TMIN_', site.name,'dynamic.pdf'))
hic.pdsi.static <- dcc(chron, TMIN, dynamic = 'static', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.static))


hic.pdsi.moving <- dcc(chron, TMIN, dynamic = 'moving', win_size = 35, win_offset = 5)

print(plot(hic.pdsi.moving))
#g_test(hic.pdsi.moving)
print(traceplot(hic.pdsi.moving))
#plot(skills(hic.pdsi.moving))
dev.off()
dev.off()
}

clim.cor(IL.clim, Hickory, 'Hickory_Grove_')
clim.cor(MNwc.clim, Bonanza, 'Bonanza_Prairie_')
clim.cor(MNwc.clim, Desoix, 'Desoix_')
clim.cor(WIsc.clim, Pleasant, 'Pleasant_Valley_Conservancy_')
clim.cor(MNec.clim, Townsend, 'Townsend_woods_')
clim.cor(MNec.clim, StCroix, 'StCroix_savanna_')
clim.cor(MNse.clim, Mound, 'Mound_prairie_')


```

